Adopting agile in the large: motivations, success factors and challenges

### 2014-05-06 journal/article.tex
Work in progress

@@

>>>>

"C: Coordinating/integrating work of many teams"

* Nightly builds with work products from all teams --> Integration problems
[P5]
The nightly build process served as the point in time when the work from five or
more Scrum teams all came together into a single installable image.  This
integration effort occurred with the production of every installable image. 
Integration across so many teams every night was the primary cause of the build
problems we faced.

* Coordination challenge --> Team goal or product goal?
* Team centric focus created strong boundaries --> Bad for coordination
[P7]
A challenge we faced in team coordination was understanding whether to optimize
the team goal or the product goal. Some teams focused too much on their own
goals and created strong boundaries, impeding coordination.

* How to manage dependencies between teams? --> No further mentionings
[P9]
Unresolved at the end of the pilot, however, was the issue of how to manage
external dependencies with other teams.

* ?
[P10]
CROSS-FUNCTIONAL DEPENDENCIES: However well we did within the team, we also
frequently ran into dependencies on other groups often in other locations –
groups such as UED (User Experience and Design), operations, legal and the
Yahoo! security team (affectionately called the Paranoids) – which were
identified as blocking issues but almost never neatly resolved within our sprint
boundaries.

* Well into the transformation there was trouble with dependencies between teams
* Management burden to manage dependencies --> Creates problem with roadmapping
* Hard to create roadmap between teams --> Teams see only their own priorities
* Reduce dependencies --> Help roadmapping
[P10]
Now  teams were  comprised of tight-knit groups that – while collocated wherever
possible – were  responsible to deliver  value regardless of location,
technology or historical allegiance.
However, within six months or so, and despite a pattern  of  success and
improvement, it became very clear that the new forking between horizontal
and vertical  teams  placed an enormous burden on the product and program
management teams to manage dependencies in a way that supported  the  broader
organizational roadmap.  Teams were structured to execute independently,  but
the broader product roadmap called for features that  implied  specific platform
improvements. Establishing a  reliable roadmap therefore required massive (and 
often unreliable) coordination between teams that were more often  than not
head-down in their own (sub-sprint) priorities.
Theoretically,  the product owners (assisted by the Music PMO) would ensure that
capabilities required by the  roadmap would appear on their own team’s backlog
at an appropriate level of priority.  In practice, it was very difficult to
identify  and track such dependencies, especially when the  feature  set  (and
associated set of dependencies) was constantly in flux.
Phase 2 of the transition (continuing today –  see Figure 4) therefore  attempts
to further reduce such dependencies and improve how they are managed over time
to improve release and roadmap planning.

* Team centricity --> Power --> However, they operate with independent priorities --> No respect for larger context (release)
* Release Framework was created
[P13]
Shortly after the initial rollout to 30+ product teams it was clear that we had
unleashed the power of our individual teams through their new found
self-organization and focus on short sprints.  However, since we aggregate the
work of many sprints into a single release it was also becoming clear that our
scrum teams were operating independently without respect for the context of the
larger release.  Herding 30+ teams working in the same codeline to a predictable
time-boxed release date is very challenging in any traditional development model
let alone an agile development model where we value self-organization over
centralized control.
Our answer to this problem was a lightweight Release Framework and the
introduction of a Release Manager for the overall product release.

* Ease coordination by maintaining code quality as top priority, above new features
[P13]
We needed to ensure that the process or framework we put into place would
promote visibility and optimization of the whole release, or of the suite of
product functionality that we deliver to our customers and our partners.
With such a large organization working on the same codeline, it was imperative
that our global priorities were consistent and clear to all scrum teams in order
to maintain quality of the release throughout the development cycle.  The
priorities were as follows:
1. Maintain existing functionality and quality of the overall codeline on a
daily basis.  All automation regression tests must pass at a 99\% passing
threshold on every checkin.
2. Building new functionality is secondary to codeline stability.
While each scrum team had to invest time to maintain the quality of the codeline
on a daily basis, they also experienced benefits from it as well.

* Individual component teams gained success --> Coordination challenges across teams
* Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
* Help: Tool support to visualize dependencies for people
[P17]
As individual component teams began to experience success, the challenge changed
to coordinating across teams. Although iterations were humming along at the team
level it was initially difficult for managers to get a complete picture of how
the release was progressing. They were unable to tell when particular features
were complete as their requirements crossed team boundaries.
Using Rally’s application for managing the development lifecycle, BMC is now
able to guide the overall release objectives with themes and high-level
features.

* Initial adopting good
* Problems with inter-team coordination --> Architecture fragile, code dialects, distrust between teams
* (Multisite, teams on different sites)
* Help: Retrospectives across teams --> Created global teams --> Shared backlog
[P24]
To embrace and implement Scrum across the different teams and organizations, we
focused on doing regular retrospectives on both team and company level.
The first couple of company retrospectives were quite hectic and required some
extra facilitation techniques to be constructive. But we did identify some of
the main problems when crossing the team boundaries and working together with
other departments.
No matter how much communication we tried to establish between the teams, we
were still far away from shared team focus on the complete product. As a result
we could see the architecture becoming more fragile, multiple code dialects and
distrust between teams were also starting to grow.
By doing regular retrospectives with all the different teams and across the
company we identified the main problems and decided it was time to create the
global Scrum teams. We wanted to improve communication and trust by working on
shared commitments between the different teams.

We still needed to improve the focus on the complete product and not on "bad
code” from other teams. The main barrier for building quality into the product
was the distrust between the teams who worked on the same product. Therefore the
two R\&D teams in Denmark and Bangladesh were merged into one global team.
Instead of two distributed teams, working with different sprint backlogs, we
established one team with one sprint backlog. The problem of building quality
into the complete product was now a shared team problem, not a management
problem.

* Scrum on team level --> Software interfaces not properly synchronized
* Problems with larger features
* Scrum on team level good --> True benefit requires entire program to be flexible for changes
[P26]
Using scrum only on the team level had led to additional problems. Software
interfaces were not properly synchronized with other development teams.
Larger features were causing synchronization problems due to dependencies in
code. Synchronization to the main waterfall program was considered problematic.
Deploying scrum on the team level had brought flexibility in the team, but the
ability to embrace the change was not truly benefiting the whole program as it
was lacking capabilities to systematically embed the changes in all teams and
levels.

* Problems with cross team integration, lack of unified build environment
[P27]
1. Our department had a lack of program-wide build scripts, build environment,
   data alignment and testing strategy.
2. Cross team integration build took too long.

* Different customers compete --> Customer centricity does not view the system as a whole
[P29]
each customer base views the system only via their process flow.
We have multiple competing customers crossing multiple co-dependent
applications, including dependencies on applications outside of our
organization; multiple funding sources; and a mixture of Waterfall and Scrum
projects.
For our software teams to be able to execute efficiently, we will need to push
the complexity and prioritization upstream, scaling teams to meet demand while
maintaining clean, simple prioritized Product Backlogs for each product.

* First cross team dependencies were reduced --> Problems with too much dependencies
* Traditional dependency management limited flexibility, which was demanded --> Inflexibility, work intensive, adversarial relationships
[P34]
From the onset, we sought to reduce cross-team dependencies as much as possible,
allowing each of the Scrum teams to “run” at their fastest pace.  However, as
the number of concurrent work streams increased, so did the number of
dependencies.  In particular, we found that adopting a SOA ran counter to our
efforts.
We quickly faced the prospect of implementing and integrating with a sizeable
set of centralized services that would be consumed by multiple project teams.
Within a few months, we had accumulated a significant number of integration
points, preventing us from delivering fully production-ready software until they
were resolved.
Accepting the fact that these dependencies would be part of the project, we
sought to create a lightweight and efficient process for resolving dependencies
and integration points.  We recognized that traditional approaches to dependency
management tend to place significant constraints on the sequence of work,
reducing the teams’ ability to independently prioritize their backlogs. 
Furthermore, highly contractual, document-centric approaches to defining
integration points are work-intensive, prone to error and can create adversarial
relationships between teams.  We wanted to establish an approach based on direct
communication and collaboration.
Our initial integration efforts involving five of the module teams of the Core
Application were reasonably successful. After some initial scrambling, the teams
self-organized and established a set of procedures for managing inter-module
integration.

* Worked decently with 5 teams and SoS --> Problems with global deployment
[P34]
However, our efforts to facilitate cross-team collaboration within the Core
Application benefited heavily from the fact that the five module teams were all
closely co-located, within the same organizational structure and employed an
active “Scrum of Scrums” model to ensure communication and coordination between
the teams.  As we pushed this approach out further across the program and the
globe, we began to experience significant challenges.

* Problem: Agile teams were thinking inside the team room walls
* Success: Thinking of the teams that will integrate to your work
* Failure: Thinking only inside the team walls
[P34]
We found that we had varying degrees of success in managing heavily-used
services.  In those cases where we were successful, it was invariably due to a
small group of people from both the consuming and providing teams taking
ownership of the services and proactively seeking out future integration points
and potential challenges.  Conversely, in those situations where such ownership
did not exist and people remained within their team walls, we invariably
experienced integration challenges: services consistently failing to meet the
consumer’s requirements and integration points that were poorly tested and
failed in system testing.

* Global effort problems --> Communication regressed to contractual --> Narrow communication channels were used: documents, e-mail
* Success: Individuals that saw effort in wide-channel communication, Individuals that thought outside team room walls
[P34]
As more and more integration occurred across the globe, we witnessed a tendency
to revert back to contractual forms of communication. As always, distributing
the development team significantly increases the challenge to directly
communicate and collaborate.  The natural tendency to rely on e-mail due to
differences in time zones frequently led to communication breakdowns and
unnecessary blockages – forming barriers to feedback. In some instances,
individuals broke through these barriers using instant messaging products, web
cams and adjusted working hours to facilitate as much direct communication as
possible.  At the same time other teams regressed, relying on documentation and
establishing an almost adversarial relationship with their integration partners.

* Large program problems --> Green builds hard to maintain
* Tools insufficient for large
[P34]
Given the size and distribution of our development team and the fact that the
program produced a single, integrated enterprise application, the complexity of
the build process and the level of effort to maintain it was significant.   As
the program grew in size, we recognized that the tool set with which we began
the project was no longer sufficient.  We began experiencing nearly continuous
broken builds due to conflation and timing issues on the build server.  At this
point, we began a lengthy initiative to replace the toolsets and revamp the
entire build process using AntHillPro and Subversion.  As a result, we created a
build process that effectively managed our myriad of dependencies and supported
a total build of the application in a reasonable time.

* Problems: Integration and dependency management
* Goal: Foster communication and collaboration instead of rigid processes
* Problems: Team avoiding external distractions, balance between team membership and project membership --> CI and green builds did not work
* Problems: Technical leads must work outside of team --> To coordinate
* Solution: Balance between own goals and project goals --> Think outside team room
* The team must think of others than them selves, too
[P34]
Over the lifetime of the project we encountered two key challenges due to the
scope and scale of this effort: integration and dependency management and
continuous integration. When addressing these challenges, we sought to adhere to
the basic principals of the Agile manifesto as much as possible: making the
delivery of working software our overriding goal and fostering communication and
collaboration instead of dictating rigid processes. We sought to empower
individuals and teams to make decisions and formulate their own processes, and
to take calculated risks without fear of reprisals.

However, we found that some of the qualities and practices that made the teams
work effectively in isolation actually reduced their effectiveness in the
large-scale environment.
For example: in smaller projects, we work aggressively to avoid external
distractions and encourage team members to focus exclusively on their individual
team goals as expressed in the backlog.  In large-scale projects, team members
must participate in cross-team activities and allocate time to project-wide
activities.  In particular, team technical leads were required to dedicate
significant portions of their time outside of the team room.  We found that many
of our team leads struggled with this task.  While they recognized the need for
cross-team communication, they resented context switching and the impact on
their time dedicated to the team’s sprint goals.
Furthermore, all of the teams struggled with balancing their identity as a team
member against their membership in the larger project. This was reflected most
keenly in our failure to sustain continuous integration and produce frequent
green builds.

Agile team must make trade-offs between stabilizing an application and
completing more story cards, the size of this project required teams to put
significantly more time into tracking down issues that were caused by other
teams or dependencies.   Many team members found this frustrating and
demoralizing.

[P43] Ryan 2008 - The Price of Agile Is Eternal Vigilance.pdf - 43:11 [We implemented a scrum-of-scru..]  (4:48-4:200)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
No memos

We 
implemented a scrum-of-scrums run by our release 
management team to identify inter-team dependencies 
and resolve merge issues before they occur.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:18 [What could not be mitigated we..]  (4:2133-4:2292)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

What could not be mitigated were issues with teams dependent on “unreliable”
Scrum teams, at least if the Scrum teams were isolated in a non-Scrum environment.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:28 [Several approaches are known t..]  (8:522-8:1054)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
No memos

Several approaches are known to co-ordinate the work of several Scrum teams working
on one product, e.. Scrum of Scrums [9] or MetaScrum [18]. The size and complexity
of SAP’s products makes it necessary to consider these techniques. To co-ordinate the
work of multiple Scrum teams and to communicate the requirements via team product
backlogs, various approaches were combined. Special product teams were formed for
this purpose. These teams can be regarded as both a permanent Scrum of Scrums and a
supervisory product owner.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:32 [Product teams were introduced ..]  (8:1076-8:1627)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Product teams were introduced as a second organizational layer above the Scrum teams.
A product team is responsible for the work of up to 7 development teams. It consists
of the product owners of those teams and the same number of team members who
are specialists in particular ﬁelds. Depending on the problem area other experts may
be included. This allows for full engineering coverage of all problem areas expected,
well-organized communication among the teams, and direct communication with the
development teams to detect and mitigate risks.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:33 [Product teams can rarely suppo..]  (9:841-9:1526)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Product teams can rarely support more than seven Scrum teams. Ifthe products or components are be too big or too complex to be developed in this organizational setup, an
intermediate organizational layer is inserted between the product team and the Scrum
teams. These teams are called “area product teams”. Their responsibility is the same as
the product teams’ responsibility with the exception of budget and ﬁnal product decisions. In this three-layer organization a productteam works similar to a project management ofﬁce. The complete budget and product responsibility is with the product team,
while the operational day-to-day decisions are handled by the area product teams.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:38 [Scaling Scrum to a multi-team ..]  (10:1223-10:1784)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [O: Large scale definition - Family: Contextual] 
No memos

Scaling Scrum to a multi-team development organization is not as easy. Scrum of
Scrums and backlog preparation for many teams can be combined by product teams
as outlined above. Looking at all known parameters of agile project management at
SAP we tend to say that not much more than 130 development employees may be
organizedin that fashion. This number sums up developers in 7 teams (max. 70 people),
the product team (max. 16), developmentinfrastructure responsibles (about 10), quality
assurance and testers (about 25), general management (about 10).

[P48] Smits 2011 - Agile Experience Report - Transition And Complexity At Cisco Voice Technology Group.pdf - 48:30 [Tracking progress of hundreds ..]  (4:1657-4:2111)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Tracking progress of hundreds of teams working on 
pieces of major features proved to be a major 
problem. Tools that we were using did not have roll 
up capability without imposing strict rules on how 
data was entered, labeled and updated. We found out 
our tools were not really scalable for the size and 
complexity of the programs we were working on. 
This proved a serious handicap in providing 
executive summary information about progress.

[P50] Sutherland 2011 - Hitting The Wall - What To Do When High Performing Scrum Teams Overwhelm Operations And Infrastructure.pdf - 50:20 [A key impediment was the distr..]  (4:1737-4:2643)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

A key impediment was the distributed 
development model in use by the Scrum teams.  
Originally, before Scrum adoption, all teams worked
together in a common environment.  There were 
many integration issues as a result and no solid 
process to rectify them.  So with the initial adoption 
of Scrum in summer 2009 it was decided that each 
individual Scrum team should have a dedicated and 
private server with integration to take place after the 
build.  In retrospect in the fall of 2009, the downside 
of the decision to break up the environment was 
being felt as integration issues were being discovered 
very late,  as  the  build  process  was  the  first  point  of 
integration.  To overcome this impediment the 
decision was made to re-integrate all teams onto a 
single  server  environment  again,  so  that  integration 
issues could be rectified before they left 
development.

[P51] Vlaanderen 2012 - Growing Into Agility - Process Implementation Paths For Scrum.pdf - 51:12 [Initially, the sprints lasted ..]  (5:230-5:717)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Initially, the sprints lasted two or three weeks, depending on
the team, causing a discrepancy between the sprint durations of some of the
teams. At the same time, interdependencies started to become more of an issue
as the amount of teams grew. This caused problems when, for instance, frontend teams needed functionality in the back-end that was not yet developed.
These problems, which consisted mainly of delays, were solved by ﬁxating sprint
duration at two weeks for all teams.

@@

* Transformation and scaling up are connected concepts. Although this research
  focuses on transformation process only, it can not be completely separated
  from discussion on scaling up, or on large scale agile implementation in
  general.
